---
title: "assaypos_7dma_model"
author: "Leah Lariscy"
date: "2023-03-17"
output: html_document
---

## Predictor and Outcome

Predictor: qPCR assay positivity

Outcome: PCR test positivity rate (daily)

## Load packages and data

```{r}
library(tidyverse)
library(tidymodels)
library(here)
```

```{r}
train_data <- readRDS(here("data/modeling_data/training_data.RDS"))
test_data <- readRDS(here("data/modeling_data/testing_data.RDS"))
```

## Define the model

The outcome is continuous, so we will use a linear regression model

```{r}
lm_mod <- linear_reg()
```

## Create recipe using training data

This is where we will set assaypos as predictor and 7dma as outcome

```{r}
recipe_7dma <- recipe(percent_pos_test ~ pos_rate_tr, data = train_data)
```

## Create workflow

```{r}
workflow_7dma <- workflow() %>% 
  add_model(lm_mod) %>% 
  add_recipe(recipe_7dma)
```

## Create model using workflow

```{r}
set.seed(626)
fit_7dma <- workflow_7dma %>% 
  fit(data = train_data)

tidy(fit_7dma)
```

## Use model on the training data

```{r}
aug_test <- augment(fit_7dma, train_data)
aug_test %>% select(percent_pos_test, .pred)
```

## Use RMSE and RSQ to evaluate model performance

```{r}
rmse <- aug_test %>% rmse(truth = percent_pos_test, .pred)
rsq <- aug_test %>% rsq(truth = percent_pos_test, .pred)
assaypos_7dma_metrics <- full_join(rmse, rsq)

saveRDS(assaypos_7dma_metrics, here("data/modeling_data/metrics/assaypos_7dma_performance.RDS"))

assaypos_7dma_metrics

```

## Plot actual value vs predicted value

```{r}
aug_test %>% ggplot(aes(percent_pos_test, .pred)) +
  geom_point() +
  stat_smooth(method = "lm")
ggsave(here("data/modeling_data/metrics/assaypos_7dma_predict.png"))
```

## Plot residual values

```{r}
aug_residual <- aug_test %>% mutate(residual = percent_pos_test - .pred)
aug_residual %>% ggplot(aes(.pred, residual)) +
  geom_point() 
  
ggsave(here("data/modeling_data/metrics/assaypos_7dma_residual.png"))
```

## Final fit on test data, then on test dates

```{r}
#testing data
fit_test_data <- workflow_7dma %>% 
  fit(data = test_data)

aug_test <- augment(fit_test_data, test_data)
aug_test %>% select(percent_pos_test, .pred)

rmse <- aug_test %>% rmse(truth = percent_pos_test, .pred)
rsq <- aug_test %>% rsq(truth = percent_pos_test, .pred)
test_metrics <- full_join(rmse, rsq)

test_metrics
```

```{r}
#testing dates
fit_test_dates <- workflow_7dma %>% 
  fit(data = test_dates)

aug_test_dates <- augment(fit_test_dates, test_dates)
aug_test_dates %>% select(percent_pos_test, .pred)

rmse <- aug_test_dates %>% rmse(truth = percent_pos_test, .pred)
rsq <- aug_test_dates %>% rsq(truth = percent_pos_test, .pred)
test_dates_metrics <- full_join(rmse, rsq)

test_dates_metrics
```

## Plot testing dates predictions vs truth

```{r}
aug_test_dates %>% ggplot(aes(x=date)) +
  stat_smooth(aes(y=percent_pos_test), method = "loess", span = 0.5, color = "black") +
  stat_smooth(aes(y=.pred), method = "loess", span = 0.5)
```
