---
title: "PP_DailyCase_model"
author: "Leah Lariscy"
date: "2023-03-17"
output: html_document
---

## Predictor and Outcome

Predictor: Percent positive technical replicates per sample

Outcome: Clinical case seven day moving average

## Load packages and data

```{r}
library(tidyverse)
library(tidymodels)
library(here)
```

```{r}
train_data <- readRDS(here("data/modeling_data/training_data.RDS"))
test_data <- readRDS(here("data/modeling_data/testing_data.RDS"))
```

## Define the model

The outcome is continuous, so we will use a linear regression model

```{r}
lm_mod <- linear_reg()
```

## Create recipe using training data

This is where we will set PP as predictor and DailyCase as outcome

```{r}
recipe_DailyCase <- recipe(pcr_pos ~ pos_rate_tr, data = train_data)
```

## Create workflow

```{r}
workflow_DailyCase <- workflow() %>% 
  add_model(lm_mod) %>% 
  add_recipe(recipe_DailyCase)
```

## Create model using workflow

```{r}
set.seed(626)
fit_DailyCase <- workflow_DailyCase %>% 
  fit(data = train_data)

tidy(fit_DailyCase)
```

## Use model on the training data

```{r}
aug_test <- augment(fit_DailyCase, train_data)
aug_test %>% select(pcr_pos, .pred)
```

## Use RMSE and RSQ to evaluate model performance

```{r}
rmse <- aug_test %>% rmse(truth = pcr_pos, .pred)
rsq <- aug_test %>% rsq(truth = pcr_pos, .pred)
PP_DailyCase_metrics <- full_join(rmse, rsq)

saveRDS(PP_DailyCase_metrics, here("data/modeling_data/metrics/PP_DailyCase_performance.RDS"))

PP_DailyCase_metrics

```

## Plot actual value vs predicted value

```{r}
aug_test %>% ggplot(aes(pcr_pos, .pred)) +
  geom_point() +
  stat_smooth(method = "lm")
ggsave(here("data/modeling_data/metrics/PP_DailyCase_predict.png"))
```

## Plot residual values

```{r}
aug_residual <- aug_test %>% mutate(residual = pcr_pos - .pred)
aug_residual %>% ggplot(aes(.pred, residual)) +
  geom_point() 
  
ggsave(here("data/modeling_data/metrics/PP_DailyCase_residual.png"))
```
